Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-21 paper.tex
describ bisim

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation. Here we omit rules 
that make the relation a congruence.
%
There are two cases where a continuation in \effCEK{C} relates to a 
continuation in \ineffCEKD. When $k \approx \kappa$, adding an identity cast on 
top of $k$ gives a related continuation. When $\kappa_1 \approx \kappa_2$, 
extending $\kappa_1$ and $\kappa_2$ with the same cast gives related results.
%
As for $v \approx v$, there is one rule for the dynamic type, and two rules for 
each ordinary value constructor.

And continuations in \ineffCEK\ that have many casts at the top correspond to 
continuations in \effCEK{C}\ where those casts are composed by $seq(c,c)$.

\begin{figure}
	\fbox{$\text{\effCEK{C}} \approx \text{\ineffCEKD}$}
	\[\dots\]

	\fbox{$\kappa \approx \kappa$}
	\begin{gather*}
	\inference{
		k \approx \kappa
	}{
		wrap(k) \approx \kappa
	}
	\quad
	\inference{
		\kappa_1 \approx \kappa_2
	}{
		ext(cast(T_1,l,T_2),\kappa_1) \approx 
		\kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
	}
	\end{gather*}

	\fbox{$v \approx v$}
	\begin{gather}
	\inference{
		v_1 \approx v_2 &
		v_2 : P
	}{
		\hcvOOinj{P}{v_1} \approx \vOOcast{v_2}{\cOOcast{P}{l}{\TOOdyn}}
	}
	\quad
	\inference{}{
		\hcvOOtt \approx \vOOtt
	}
	\quad
	\inference{}{
		\hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
	}
	\\
	\inference{
		\rho_1 \approx \rho_2 &
		dom(\rho_1) = dom(\rho_2) = \Gamma \\
		\Gamma , x : T_1 \vdash e : T_2
	}{
		\hcvOOfun{id(T_1)}{\rho_1}{x}{e}{id(T_2)}
		\approx
		\vOOfun{\rho_2}{x}{e}
	}
	\quad
	\inference{
		v_1 \approx v_2 & v_3 \approx v_4 &
		\judgeType{v_1}{T_1} &
		\judgeType{v_3}{T_2}
	}{
		\hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
		\approx
		\vOOcons{v_2}{v_3}
	}
	\\
	\inference{
		\hcvOOfun{c_1}{\rho}{x}{e}{c_2} \approx v_2 \\
%		\judgeType{\hcvOOfun{c_1}{\rho}{x}{e}{c_2}}{\POOfun{T_1}{T_2}} \\
		\judgeType{v_2}{\POOfun{T_1}{T_2}} \\
		c_1' = seq(cast(T_3,l,T1),c_1) \\
		c_2' = seq(c_2,cast(T_2,l,T_4))
	}{
		\hcvOOfun{c_1'}{\rho}{x}{e}{c_2'}
		\approx
		\vOOcast{v}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
	}
	\quad
	\inference{
		\hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 \\
		\judgeType{v_2}{\POOprod{T_1}{T_2}} \\
		c_1' = seq(c_1,cast(T_1,l,T_3)) \\
		c_2' = seq(c_2,cast(T_2,l,T_4))
	}{
		\hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
		\vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
	}
	\\
	\inference{
		v_1 \approx v_2 & v_1 : T
	}{
		\hcvOOinl{v_1}{id(T)} \approx \vOOinl{v_2}
	}
	\quad
	\inference{
		\hcvOOinl{v_1}{c} \approx v_2 &
		v_2 : \POOsum{T_1}{T_2} &
		c' = seq(c,cast(T_1,l,T_3))
	}{
		\hcvOOinl{v_1}{c'} \approx
		\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
	}
	\\
	\inference{
	v_1 \approx v_2 & v_1 : T
	}{
	\hcvOOinr{v_1}{id(T)} \approx \vOOinr{v_2}
	}
	\quad
	\inference{
	\hcvOOinr{v_1}{c} \approx v_2 &
	v_2 : \POOsum{T_1}{T_2} &
	c' = seq(c,cast(T_2,l,T_4))
	}{
	\hcvOOinr{v_1}{c'} \approx
	\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
	}
	\end{gather}
	\caption{Bisimulation between \effCEK{C} and \ineffCEKD}
	\label{fig:bisim-SC-D}
\end{figure}

The bisimulation relation between states ($\eqvSD{C}{s}{s}$) is mostly derived 
from the state definition: \begin{gather*}
\inference{
	\rho_1 \approx \rho_2 &
	\kappa_1 \approx \kappa_2
}{
	\sOOinspect{e}{\rho_1}{\kappa_1} \approx \sOOinspect{e}{\rho_2}{\kappa_2}
}
\quad
\inference{
	\kappa_1 \approx \kappa_2 &
	v_1 \approx v_2
}{
	\sOOreturn{v_1}{\kappa_1} \approx \sOOreturn{v_2}{\kappa_2}
}
\quad
\inference{
}{
	\sOOhalt{o} \approx \sOOhalt{o}
}
\end{gather*}

The bisimulation relation between environments and cast results are derived 
from their definitions as well. And the relation from continuations in 
\ineffCEKD\ and the pre-continuation in \effCEK{C} is in the natural way, for 
example: \begin{gather*}
\inference{
	\rho_1 \approx \rho_2 &
	\kappa_1 \approx \kappa_2
}{
	\kOOconsI{e}{\rho_1}{\kappa_1} \approx \kOOconsI{e}{\rho_2}{\kappa_2}
}
\end{gather*}

And the relation between continuations is defined in terms of $wrap(k)$ and 
$ext(c,k) $: \begin{gather*}
\inference{
	k \approx \kappa
}{
	wrap(k) \approx \kappa
}
\quad
\inference{
	\kappa_1 \approx \kappa_2
}{
	ext(cast(T_1,l,T_2),\kappa_1)
	\approx
	\kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2} 
}
\end{gather*}

The relation between values is more special. When the \ineffCEKD\ value is 
non-casted, the casts in the \effCEK{C} value are identity. Otherwise, we push 
the cast into the value constructors, except injections. For examples: 
\begin{gather*}
\inference{
	v_1, v_2 : P &
	v_1 \approx v_2
}{
	\hcvOOinj{P_1}{v_1} \approx \vOOcast{v_2}{\cOOcast{P}{l}{\TOOdyn}}
}
\quad
\inference{
	v_1, v_2 : T &
	v_1 \approx v_2
}{
	\hcvOOinl{v_1}{id(T)} \approx \vOOinl{v_2}
}
\\
\inference{
	(\hcvOOinl{v}{c}, v_2) : \POOsum{T_1}{T_2}
}{
	\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))} \approx 
	\vOOcast{v_2}{\cOOcast{T_1}{l}{T_3}}
}
%\quad
%\inference{
%	\Gamma , x : T_1 \vdash T_2 &
%	\rho_1 \approx \rho_2
%}{
%	\hcvOOfun{id(T_1)}{\rho_1}{x}{e}{id(T_2)} \approx \vOOfun{\rho_2}{x}{e}
%}
%\quad
%\inference{
%	\hcvOOcons{v_1}{c_1}{v_2}{c_2}
%	\approx
%	v
%}{
%	\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))}
%}
%\\
%\inference{
%	\hcvOOfun{c_1}{\rho}{x}{e}{c_2} : \POOfun{T_1}{T_2} &
%	\hcvOOfun{c_1}{\rho}{x}{e}{c_2} \approx v
%}{
%	\hcvOOfun{seq(cast(T_3,l,T_1), c_1)}{\rho}{x}{e}{seq(c_2,cast(T_2,l,T_4))}
%	\approx
%	\vOOcast{v}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T3}{T_4}}}
%}
\end{gather*}

>>>>

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation. Here we omit rules 
that make the relation a congruence.
%
The relation $\kappa \approx \kappa$ relates two continuations. 
%
If $k \approx \kappa$, adding an identity cast on top of $k$ gives a related 
continuation. 
%
If $\kappa_1 \approx \kappa_2$, extending $\kappa_1$ and $\kappa_2$ with the 
same cast gives related results.
%
%
The relation $v \approx v : T$ relates two value and also report their type. We 
will see shortly why reporting the type. 
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
Unit values are of course related.
%
To relate to a non-casted function in \ineffCEKD, the casts in the \effCEK{C} 
function must be identity.
%
If functions are related, we can apply the same cast on both side.

\todo[inline]{In Fig.~\ref{fig:bisim-SC-D}, how to suppress rule numbers like 
(1), (2), etc.? -KC}
\begin{figure}
%	\fbox{$\text{\effCEK{C}} \approx \text{\ineffCEKD}$}
	\[\dots\]

	\fbox{$\kappa \approx \kappa$}
	\begin{gather*}
	\inference{
		k \approx \kappa
	}{
		wrap(k) \approx \kappa
	}
	\quad
	\inference{
		\kappa_1 \approx \kappa_2
	}{
		ext(cast(T_1,l,T_2),\kappa_1) \approx 
		\kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
	}
\\
	\end{gather*}
\\	
	\fbox{$v \approx v : T$}
	\begin{gather}
	\inference{
		v_1 \approx v_2 : P
	}{
		\hcvOOinj{P}{v_1} \approx \vOOcast{v_2}{\cOOcast{P}{l}{\TOOdyn}}
		: \TOOdyn
	}
\\ \\
	\inference{}{
		\hcvOOtt \approx \vOOtt : \POOunit
	}
	\quad
	\inference{}{
		\hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
		: \POOunit
	}
	\\ \\
	\inference{
		\rho_1 \approx \rho_2 &
		dom(\rho_1) = dom(\rho_2) = \Gamma \\
		\Gamma , x : T_1 \vdash e : T_2
	}{
		\hcvOOfun{id(T_1)}{\rho_1}{x}{e}{id(T_2)}
		\approx
		\vOOfun{\rho_2}{x}{e}
		: \POOfun{T_1}{T_2}
	}
\\ \\
\inference{
\hcvOOfun{c_1}{\rho}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
%		\judgeType{\hcvOOfun{c_1}{\rho}{x}{e}{c_2}}{\POOfun{T_1}{T_2}} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\rho}{x}{e}{c_2'}
\approx
\vOOcast{v}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
	\inference{
		v_1 \approx v_2 : T_1 &
		v_3 \approx v_4 : T_2 &
	}{
		\hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
		\approx
		\vOOcons{v_2}{v_3}
		: \POOprod{T_1}{T_2}
	}
\\ \\
	\inference{
		\hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOsum{T_1}{T_2} \\
		c_1' = seq(c_1,cast(T_1,l,T_3)) \\
		c_2' = seq(c_2,cast(T_2,l,T_4))
	}{
		\hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
		\vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
		: \POOprod{T_3}{T_4}
	}
\\ \\
	\inference{
		v_1 \approx v_2 : T
	}{
		\hcvOOinl{v_1}{id(T)} \approx \vOOinl{v_2}
		: \POOsum{T}{T'}
	}
	\quad
	\inference{
		\hcvOOinl{v_1}{c} \approx v_2
		: \POOsum{T_1}{T_2} &
		c' = seq(c,cast(T_1,l,T_3))
	}{
		\hcvOOinl{v_1}{c'} \approx
		\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
		: \POOsum{T}{T'}	
	}
	\\
	\inference{
	v_1 \approx v_2 : T
	}{
	\hcvOOinr{v_1}{id(T)} \approx \vOOinr{v_2}
	: \POOsum{T'}{T}
	}
	\quad
	\inference{
	\hcvOOinr{v_1}{c} \approx v_2
	: \POOsum{T_1}{T_2} &
	c' = seq(c,cast(T_2,l,T_4))
	}{
	\hcvOOinr{v_1}{c'} \approx
	\vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
	: \POOsum{T'}{T}
	}
	\end{gather}
	\caption{Bisimulation between \effCEK{C} and \ineffCEKD}
	\label{fig:bisim-SC-D}
\end{figure}
